Priors for Latent Hawkes

### 2020-06-25 output_scripts/New_Priors.Rmd
exploring fits

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,value))  + geom_point()  +
    theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

@@

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    #rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    #pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    #mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,mean_rank))  + geom_point()  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

@@

  mean_rank_orig <- apply(sim_cohort_mmhp$f, 2, mean)
  mean_values <- tibble(mean_rank_dc, mean_rank_orig,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    pivot_longer(cols = c(mean_rank_dc,mean_rank_orig),names_to = "model") %>%
    mutate(model = as.factor(model)) %>%  
    ggplot(aes(isi,value))  + 
    geom_jitter(aes(color=model), alpha = 0.5,width = 0.1)  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("Cohort",current_cohort,sep=""))
  print(p1)

@@

  m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))
  indep_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/immhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  matrix_lst_plot <- list( indep_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                           m2_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
    m3_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                          m3_pr_dc[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]]
                          ) 
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-150,max=150,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("I-MMHP","DC-HP","C-MMHP","DC-C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                    matrix.mar=c(2.5,1.5,2.5,1)
                    )
}
```

>>>>

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,value))  + geom_point()  +
    theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

@@

  mean_values <- tibble(mean_rank,mean_gamma,mean_zeta,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    #rename(gamma_out = mean_gamma,zeta_in = mean_zeta) %>%
    #pivot_longer(cols = c(gamma_out,zeta_in),names_to = "out_in" ) %>%
    #mutate(out_in = as.factor(out_in)) %>%
    ggplot(aes(isi,mean_rank))  + geom_point()  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

@@

  mean_rank_orig <- apply(sim_cohort_mmhp$f, 2, mean)
  mean_values <- tibble(mean_rank_dc, mean_rank_orig,
                        isi = expert_rank_10[[current_cohort]])
  p1 <- mean_values %>%
    pivot_longer(cols = c(mean_rank_dc,mean_rank_orig),names_to = "model") %>%
    mutate(model = as.factor(model)) %>%  
    ggplot(aes(isi,value))  + 
    geom_jitter(aes(color=model), alpha = 0.5,width = 0.1)  +
    # theme(legend.title = element_blank()) + facet_wrap(~out_in,scales="free") +
    scale_x_continuous(breaks = 1:12) + xlab("I&SI Ranking") +
    ggtitle(paste("DC C-MMHP Cohort",current_cohort,sep=""))
  print(p1)

@@

  m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                     "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))
  m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                     "/chp_pr_matrix_",cohort_names[current_cohort],
                     ".RDS",sep=''))

  matrix_lst_plot <- list( m1_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                           m2_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
    m3_pr[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]],
                          m3_pr_dc[rev(expert_rank_10[[current_cohort]]),
                                              expert_rank_10[[current_cohort]]]
                          ) 
  myMultiMatrixPlot(X=matrix_lst_plot,no_matrix=4,n_row=2,
                    xLabels=expert_rank_10[[current_cohort]],
                    yLabels=rev(expert_rank_10[[current_cohort]]),
                    min=-150,max=150,axis_cex=2,title_cex = 1.8,
                    colorPalette="RdBu",if.emp=FALSE,
                    #legend.mar=c(0.5,.5,0.5,0.5),
                    title_lst=list("C-HP","DC-HP","C-MMHP","DC-C-MMHP"), 
                    #title_lst = list("I-MMHP"),
                    col_axis=c(-150,-75,0,75,150), fake_matrix=FALSE,
                    matrix.mar=c(2.5,1.5,2.5,1)
                    )
}
```

# Investigate the Pearson Residuals further

m3_pr_dc <- readRDS(paste(dc_data_path,cohort_names[current_cohort],
                   "/cmmhp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))
m1_pr <- readRDS(paste(save_data_path,cohort_names[current_cohort],
                   "/chp_pr_matrix_",cohort_names[current_cohort],
                   ".RDS",sep=''))

max(m3_pr_dc)

which(m3_pr_dc == max(m3_pr_dc),arr.ind = T)

i <- 10
j <- 1

```
Let's examine this pair further.

stan_input_lst <- prepareDataStan(current_cohort)
row_ind <- which(stan_input_lst$I_fit == i & stan_input_lst$J_fit == j)

pair_events <- stan_input_lst$event_matrix[row_ind, ,]
window_lengths <- stan_input_lst$finishing_time

# this is the events between this pair

fit_cohort_mmhp_dc <- fit_cohort_mmhp
sim_cohort_mmhp_dc <- sim_cohort_mmhp

load(paste(save_data_path,cohort_names[current_cohort],
         "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
         ".RData",sep=''))

lambda0_pair <- mean(sim_cohort_mmhp_dc$gamma[,i]) +
  mean(sim_cohort_mmhp_dc$zeta[,j])

lambda0_pair

mean(sim_cohort_mmhp$lambda0)

mean(sim_cohort_mmhp_dc$f[,i]) - mean(sim_cohort_mmhp$f[,i])

mean(sim_cohort_mmhp_dc$f[,j]) - mean(sim_cohort_mmhp$f[,j])

# essentially no difference in the latent ranks, some differences in other parameters

mean(sim_cohort_mmhp_dc$beta) - mean(sim_cohort_mmhp$beta)
mean(sim_cohort_mmhp_dc$eta_1) - mean(sim_cohort_mmhp$eta_1)
mean(sim_cohort_mmhp_dc$eta_2) - mean(sim_cohort_mmhp$eta_2)
mean(sim_cohort_mmhp_dc$eta_3) - mean(sim_cohort_mmhp$eta_3)

```{r}
pair_events <- as.data.frame(t(pair_events))
colnames(pair_events) <- paste("wind_",c(1:ncol(pair_events)),sep = "")

tidy_events <- tibble(pair_events) %>% 
  pivot_longer(starts_with("wind"), names_to = "wind") %>%
  filter(value > 0) %>%
  group_by(wind) %>%
  summarise(times = list(value)) %>%
  mutate(wind_num = stringr::str_extract(wind, pattern = "[:digit:]+"),
         wind_num = as.numeric(wind_num)) %>%
  arrange(wind_num)

tidy_events # this is only windows with events now

windows_pair <- tibble(wind_num = 1:length(window_lengths),window_lengths)

events <- windows_pair %>% left_join(tidy_events) %>% select(-wind) %>%
  rowwise() %>%
  mutate(num_events = length(times))

Then we need to try get the latent states for each of these windows also.

```{r}
load(paste(dc_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[70]]
rm(interpolation_array_list) # too big to keep around

However, the only windows with any uncertainty in their latent state, for this pair, are those where

mean_state <- matrix(nrow =  length(window_lengths), ncol = no_segments)

for(i in 1:length(window_lengths)) {
  mean_state[i, ] <- apply(latent_pair[[i]], 1,var)
}

as.data.frame(mean_state)

Did this happen when not using the degree correction? It seems likely.

```{r}
load(paste(save_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[row_ind]]
rm(interpolation_array_list) # too big to keep around

mean_state <- matrix(nrow =  length(window_lengths), ncol = no_segments)

for(i in 1:length(window_lengths)) {
  mean_state[i, ] <- apply(latent_pair[[i]], 1,mean)
}

as.data.frame(mean_state)
```

```{r}
tibble(send = stan_input_lst$I_fit, rec = stan_input_lst$J_fit,
       num_events = apply(stan_input_lst$Nm, 1, sum), id = row_number(stan_input_lst$I_fit)) %>%
  arrange(desc(num_events))

Let's check from before I updated to account for empty windows.

This is still to big to load locally.

```{r, eval=FALSE}
old_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/extend_stan/"

current_cohort <- 1

inter_array_list <- readRDS(paste(old_path,cohort_names[current_cohort],
                                  "/cohort_mmhp_state_array_list_",
                                  cohort_names[current_cohort],".RDS",sep = ""))

pair_array <- inter_array_list[[row_ind]]
rm(inter_array_list)

dim(pair_array[[27]])

```
Let's look at a single pair from Jing's original fit.

```{r}
single_pair <- readRDS("C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/single_inter_array.RDS")

dim(single_pair[[2]])
apply(single_pair[[25]],1,var)

for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(single_pair[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(single_pair[[i]],1,var)),"\n",sep = ""))
}

```
We can repeat this for the previous fit also and compare.

```{r}
for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(pair_array[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(pair_array[[i]],1,var)),"\n",sep = ""))
}

```
These numbers make no sense for sure.

Let's compare to the current model.

```{r}
load(paste(dc_data_path,cohort_names[current_cohort],
         "/cmmhp_est_zt_",cohort_names[current_cohort],
         ".RData",sep=''))

latent_pair <- interpolation_array_list[[69]]

for(i in 1:27){
  try(cat("Mean for window ",i, " is ",
          mean(apply(latent_pair[[i]],1,mean)), "\n",sep=""))
  try(cat("Var for window ",i, " is ",
          mean(apply(latent_pair[[i]],1,var)),"\n",sep = ""))
}

```
So a handful of windows give different results but overall there is little 
difference, and it seems likely all of these differences are due to different
parameters and also running less segments for each window.
